---
title: "Penguin size comparisons"
author: "Lory Salazar"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

## Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap penguins

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g) # pull out values for variable we specify

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g) 
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?
(this is a two-sided question)
```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m) # to get effect size
```

Are male chinstrap penguins significantly larger in body mass than female chinstrap penguins?
(this is a one-sided question)

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # greater or less depending on the direction of your question. it should line up with the order you put the vectors so that this code asks are chinstraps_m greater than chinstraps_f
```

## One-way ANOVA

### 1. Table of counts by species and sex

```{r}
penguin_counts <- penguins %>% 
  count(species, sex) # gives the n for the observations of the variables you listed

penguin_counts
```

Create a subset of only female penguins
```{r}
penguin_f <- penguins %>% 
  filter(sex == "female")

ggplot(data = penguin_f, aes(x = flipper_length_mm)) +
  geom_histogram(aes(fill = species)) + # on its own this line would split the species by color in one graph
  facet_wrap(~species, scales = "free") # splits into different plots, scales changes so they each have diff scales

ggplot(data = penguin_f, aes(sample = flipper_length_mm)) +
  geom_qq(aes(color = species)) + # change to color (not fill) because qqplot gives pints
  facet_wrap(~species, scales = "free")
```

### Summary statistics

Starting from penguin_f, create a summary table that groups by species and returns the following:

- mean of flipper length
- sd ""
- sample size
- variance
- standard error = sd / sqrt(n)

```{r}
penguin_f_summary <- penguin_f %>% 
  group_by(species) %>% 
  summarize(
    mean_flipper_length = mean(flipper_length_mm),
    sd_flipper_length = sd(flipper_length_mm),
    sample_size = n(),
    se_flipper_length = sd(flipper_length_mm)/sqrt(n()), # to find standard error using the equation
    var_flipper_length = var(flipper_length_mm)
  )

leveneTest(flipper_length_mm ~ species, data = penguin_f) # looks at variance in flipper length as a function of species
```
Null hypothesis here is that the  variance is equal
For this test, the p-value is above the significance level so null hypothesis is retained. We have satisfied the equal variance condition for an ANOVA test. 

### Beeswarm plot with mean +- 1sd

```{r}
ggplot() +
  geom_beeswarm(data = penguin_f, 
                aes(x = species, y = flipper_length_mm),
                color = "gray70",
                size = 3,
                alpha = 0.5) +
  geom_point(data = penguin_f_summary,
             aes(x = species, y = mean_flipper_length),
             size = 4) +
  geom_errorbar(data = penguin_f_summary,
                aes(x = species,
                    ymax = mean_flipper_length + sd_flipper_length, # assigning whiskers to extend +1 and -1 sd
                    ymin = mean_flipper_length - sd_flipper_length),
                width = 0.1) + # adjusts the whisker size
  theme_light()
```

### One-way ANOVA
aov in R default runs type I ANOVA

```{r}
penguin_aov <- aov(flipper_length_mm ~ species, data = penguin_f)
summary(penguin_aov) 

TukeyHSD(penguin_aov) # after ANOVA, only if we have a very small p-value
```

Type III ANOVA with the car package:
```{r}
penguin_lm <- lm(flipper_length_mm ~ species, data = penguin_f) # creating a linear model
penguin_aov_t3 <- Anova(penguin_lm, type = "III") # for a type III ANOVA (if you have a very unbalanced design)
penguin_aov_t3
```
Similar f value to type I; this usually will happen. Will draw the same conclusion from before.
