---
title: "Penguin size comparisons"
author: "Lory Salazar"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(palmerpenguins)
library(effsize)
library(broom)
library(ggbeeswarm)
library(car)
```

# Part 1. T-tests

Mass comparison: male & female chinstrap penguins

```{r}
chinstraps <- penguins %>% 
  filter(species == "Chinstrap")

## Histograms
ggplot(data = chinstraps, aes(x = body_mass_g)) +
  geom_histogram(bins = 10) +
  facet_wrap(~sex)

## QQ plot
ggplot(data = chinstraps, aes(sample = body_mass_g)) +
  geom_qq() +
  facet_wrap(~sex)

## Get descriptive statistics
chinstraps %>% 
  group_by(sex) %>% 
  summarize(
    mean_mass = mean(body_mass_g),
    median_mass = median(body_mass_g),
    sd_mass = sd(body_mass_g),
    sample_size = n()
  )
```

Get vectors of body_mass_g for female and male chinstrap penguins

```{r}
chinstraps_f <- chinstraps %>% 
  filter(sex == "female") %>% 
  pull(body_mass_g) # pull out values for variable we specify

chinstraps_m <- chinstraps %>% 
  filter(sex == "male") %>% 
  pull(body_mass_g) 
```

### T-tests, a couple different ways

1. Is there a significant difference in body mass between male and female chinstrap penguins?
(this is a two-sided question)
```{r}
chinstrap_ttest <- t.test(chinstraps_f, chinstraps_m)
chinstrap_ttest

cohen.d(chinstraps_f, chinstraps_m) # to get effect size
```

Are male chinstrap penguins significantly larger in body mass than female chinstrap penguins?
(this is a one-sided question)

```{r}
t.test(chinstraps_m, chinstraps_f, alternative = "greater") # greater or less depending on the direction of your question. it should line up with the order you put the vectors so that this code asks are chinstraps_m greater than chinstraps_f
```

